# Numeric vectors

**Learning objectives:**

- Create numeric vector 
    * `c` , `rep`/`seq` and `:`, 
    * (pseudo)random numbers, `sample()` / `r`+`unif`/`norm`.   
    * reading with `scan`
    
- Naming our first object: `<-` 

- **Vectorized** mathematical functions 
    * Some ubiquitous operations: `abs`, `round`, `exp`. 
    * Probability distribution: continous (uniform) / discrete (Poison). 
    * Arithmetic operations
        + **Vectorized** arithmetic operators. 
        + **Recycling rule**  
        + Operator precedence 
        + shortly accumulating and aggregating 

## Creating numeric vector 1 {-}

### Numeric constant

```{r}
-3.14
1.23e3
NA_real_
typeof(NA_real_)
```

### Vectors 

```{r}
c(1, 5, 9)
# c( ... ) dot-dot-dot are called ellipsis: any arbitrary number of arguments
rep(1, 5)
# rep(x, times = 1, length.out = NA, each = 1)
```

R allows positional matching of argument 

```{r}
rep(c(1, 2, 3), 4)   # positional matching of arguments: `x`, then `times`
rep(c(1, 2, 3), times=4)    # `times` is the second argument
rep(x=c(1, 2, 3), times=4)  # keyword arguments of the form name=value
rep(times=4, x=c(1, 2, 3))  # keyword arguments can be given in any order
rep(times=4, c(1, 2, 3))    # mixed positional and keyword arguments
```

You should avoid partial matching! 

### Arithmetic progressions `seq` and `:`

```{r}
seq(1, 15, 2) 
seq(from = 1, to = 15, by = 2)
seq(to = 15, by = 2)
```

`length.out` argument can also be used 

```{r}
1:10
-1:10
-1:-10
-(1:10)
-(1):10
```

### Generating pseudorandom numbers

```{r}
runif(2) # uniform U(min = 0, max = 1)
rnorm(10) # normal N(mean = 0, sd = 1)
```

We are going to see a list of them later!

`Sample` to sample item from a given vector:

```{r}
sample(1:10, 5, replace = TRUE)
sample(10, 5, replace = TRUE)
```

You can use `set.seed(42)` to specify a state for the Random Number Generator (see `help(RNG)`) 

### reading data with `scan` 

```{bash}
head data/euraud-20200101-20200630.csv
```

We can use `scan`:

```{r}
scan(paste0("https://github.com/gagolews/teaching-data/raw/",
    "master/marek/euraud-20200101-20200630.csv"), comment.char = "#") 
```

```{r}
scan("data/euraud-20200101-20200630.csv", comment.char = "#")
```

What does the following arguments (use cases)? 

- `dec`  
- `sep`  
- `what` 
- `na.strings` 

## Creating named objects {-}


## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
