# Numeric vectors

**Learning objectives:**

- Create numeric vector 
    * `c` , `rep`/`seq` and `:`, 
    * (pseudo)random numbers, `sample()` / `r`+`unif`/`norm`.   
    * reading with `scan`
    
- Naming our first object: `<-` 

- **Vectorized** mathematical functions 
    * Some ubiquitous operations: `abs`, `round`, `exp`. 
    * Probability distribution: continous (uniform) / discrete (Poison). 
    * Arithmetic operations
        + **Vectorized** arithmetic operators. 
        + **Recycling rule**  
        + Operator precedence 
        + shortly accumulating and aggregating 

## Creating numeric vector 1 {-}

### Numeric constant

```{r}
-3.14
1.23e3
NA_real_
typeof(NA_real_)
```

### Vectors 

```{r}
c(1, 5, 9)
# c( ... ) dot-dot-dot are called ellipsis: any arbitrary number of arguments
rep(1, 5)
# rep(x, times = 1, length.out = NA, each = 1)
```

R allows positional matching of argument 

```{r}
rep(c(1, 2, 3), 4)   # positional matching of arguments: `x`, then `times`
rep(c(1, 2, 3), times=4)    # `times` is the second argument
rep(x=c(1, 2, 3), times=4)  # keyword arguments of the form name=value
rep(times=4, x=c(1, 2, 3))  # keyword arguments can be given in any order
rep(times=4, c(1, 2, 3))    # mixed positional and keyword arguments
```

You should avoid partial matching! 

### Arithmetic progressions `seq` and `:`

```{r}
seq(1, 15, 2) 
seq(from = 1, to = 15, by = 2)
seq(to = 15, by = 2)
```

`length.out` argument can also be used 

```{r}
1:10
-1:10
-1:-10
-(1:10)
-(1):10
```

### Generating pseudorandom numbers

```{r}
runif(2) # uniform U(min = 0, max = 1)
rnorm(10) # normal N(mean = 0, sd = 1)
```

We are going to see a list of them later!

`Sample` to sample items from a given vector:

```{r}
sample(1:10, 5, replace = TRUE)
sample(10, 5, replace = TRUE)
```

You can use `set.seed(42)` to specify a state for the Random Number Generator (see `help(RNG)`) 

### reading data with `scan` 

```{bash}
head data/euraud-20200101-20200630.csv
```

We can use `scan`:

```{r}
scan(paste0("https://github.com/gagolews/teaching-data/raw/",
    "master/marek/euraud-20200101-20200630.csv"), comment.char = "#") 
```

```{r}
scan("data/euraud-20200101-20200630.csv", comment.char = "#")
```

What does the following arguments (use cases)? 

- `dec`  
- `sep`  
- `what` 
- `na.strings` 

## Creating named objects {-}

- You can assign (memorized) and object with `<-`

- Name are case-sensitive: `bob` != `Bob` 

- `.` is legal but not if followed by number and starting with it

- `'if`, `for`, `function`, `next` `TRUE` are reserved


You should follow a naming conventions (usually follow project guidelines). 

For naming temporary object: 

- vectors: x, y z

- matrices: A, B, ... , X, Y, Z

- integer indexes: i, j , k 

- object size: n , m, p or nx, ny, etc 

One tips: 

```{r}
(x <- 1:3) # allow you to print the object
```


## Vectorised mathematical functions {-}

$$\textbf{x} \quad of \quad length \quad n  \quad (x_1, x_2, ..., x_n)$$

`abs()` is vectorized: 

```{r}
abs(c(2, 0, -1, -3, NA_real_))
```

Instead of defined to act on a single value ("scalar") it is applied on each elements: 

$$ |\textbf{x}| = (|x_1|, |x_2|, ..., |x_n|) $$

It is the same for `round`, `floor`, etc but also `log` and `exp` (keep it mind changing the base of the log is an argument, default is `base = exp(1)`) 

## Probability distributions: 

R has a lot of univariate distributions both discrete and continuous. 

I use very frequently `*unif`, `*norm` for continuous and `*binon`, `*pois` for discrete. 
For continuous distributions `*` can be:

- `d` : probability density function (PDF)

- `p` : cumulative distribution function (CDF)

- `q` : quantile function (1/CDF)

- `r` : generate random number (deviates?)

For discretes distributions: 

- `p` and `r` are the sam as above

- `d` give you the probability mass function (PMF)

- `q` is also the quantile function but defined as a generalised 1/CDF

You can find packages that implement other commons functions if needed. 


## Arithmetic operations {-}

### Vectorised arithmetic operators

R have the classic operators (`+`, `-`, `*`, `/`, `%/%`, `%%`, `^` == `**`) but they are vectorised: 

```{r}
c(1, 2, 3) * c(10, 100, 1000)

`*`(c(1, 2, 3), c(10, 100, 1000))
```

### Recycling rule 

`+` & co are called operators and `c(1, 2, 3)` is an operands. 

When operands have different length they are recycled: 

```{r}
0:7 /3
1:10 * c(-1, 1)
2 ^ (0:10)
```


But what happens if one operands can't be recycled in its entirety?

```{r}
c(1, 10 , 100) * 1:8
```

Most of the time it is used for vector scalar operations but it can be used in usefull in other schemes. 

`pmin` and `pmax` have similar behavior: 

```{r}
pmin(3, 1:5)
```

### Operator precedence

Rules that govern the order of computation: 

I prefer them in from highest to lowest precedence:

- `^` 

- `-` and `+` (unary)

- `:` 

- `%%` and `%/%`

- `*` and `/`

- `-` and `+` (binary)

- `<-`

Usually left to right except `^` and `<-`

In doubt you can always use some bracket `()` (it makes also the code more readable) !

### Accumulating 

Sometimes you do not want element-wise operations: 

You can then use the `cumsum()`/ `cumprod()` family of functions. 

```{r}
cumsum(1:8)
cummin(c(3, 2, 4, 5, 1, 6, 0))
```

`diff` is very interesting function that return lagged differences

### Aggregating 

Sometimes we just want the last "cumulants" that summarise all the imputs. 

We can then use `sum()`, `prod()`, `min()`. 

You also have `mean()`, `var()` and `sd()` that are very hany 

Note `var()` is `sum((x- mean(x))^2) / (length(x) - 1)`

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
