# Vector indexing

## Basics
### `head()` and `tail()`
+ `head()` and `tail()` return the first or last six elements of an atomic vector
```{r}
x <- 1:10
head(x)
```

```{r}
tail(x)
```

## `'['` - Extract or Replace
+ We can index with the familiar `'['`
+ It's an actual R function!
```{r}
base::'['
```
+ `'['` behaves differently for atomic vectors, lists, data.frames and many other types
```{r}
length(methods('['))
```
+ It has methods for tons of different object classes!

## `'['` - Extract or Replace
```{r}
x <- seq(10, 100, 10)
# First element
x[1]
# Last element
x[length(x)]
# First three elements
x[1:3]
```
## `'['` - Extract or Replace
+ `'['` takes both numeric and logical vectors
+ Therefore, we can index with an expression or the `which()` function
```{r}
x <- runif(10, 1, 10)
y <- runif(10, 1, 10)

# These are the same
all.equal(x[x>y], # indexer returns logical vector
         x[which(x>y)]) # indexer returns numeric vector
```

## `'['` - Extract or Replace
+ Note: `which()` by default removes `NA` values, while `'['` does not.
## `'['` - Extract or Replace
+ To *exclude* elements, we can pass negative integers to `'['`
```{r}
x[-1]
```
+ E.g.
```{r}
all.equal(x[-1], x[2:10])
```
## `'['` - Extract or Replace
+ `'['` returns `NA` when we pass `NA` to it
```{r}
x[c(NA, 2, 4)]
```
+ For lists, `'['` returns `NULL` when we pass `NA`

## `'['` - Extract or Replace
+ For named vectors, we can pass characters to `'['`
```{r}
x <- structure(x, names=letters[1:10])
x[c("a", "f", "a", "g", "z")]
```
## `'['` - Extract or Replace
+ `'['` works on lists, and always returns a list
```{r}
y <- list(a = 1:10,
          b = 11:20)

y[c(1,2)]
```

## `'[['` and recursiveness
+ `'[['` can be used to select a single element, removing the names attribute.
+ For lists, `'[['` does not return a list but an object of the class of the underlying list element
```{r}
x <- list(a = c(1,2,3),
          b = c(4, 5, 6))
data.class(x[1])
```

```{r}
# vs.
data.class(x[[1]])
```
## `'[['` and recursiveness
+ `'[['` operates recursively when given mulitple indexes.
+ Compare:
```{r}
# atomic vector with '['
x <- 1:10
x[c(1, 2)]
```

```{r}
y <- list(a = 1:10,
          b = 11:20)

# list with '[['
y[[c(1,2)]]   # equivalent to y[[1]][[2]]
```
## `'[<-'` and `'[[<-'`
+ The "replacement versions" of these functions are `'[<-'` and `'[[<-'`
+ These are actually separate functions, that allow us to assign values at a specific index (see Chapter 9)

```{r}
y[[c(1,2)]] <- 999
y
```
## Useful functions

## `match()` and `%in%`
+ `match()` returns the indices of elements in `x` that have matching values in `table` 
```{r}
table <- sample(c(1:10, NA), size = 100, replace = TRUE)
x <- c(5, 6, NA)
# returns indices in table
match(x, table, nomatch = 0)
```

+ `%in%` is an intuitive wrapper that returns a logical vector instead

```{r}
# returns logical in x
x %in% table
```
## `findInterval()`

+ `findInterval` returns the interval range in `interval` that contains each element in `x`

```{r}
x <- runif(10)

# a vector of length 4 has 3 intervals
interval <- c(-Inf, .25, .5, Inf)

findInterval(x, interval)
```
## `findInterval() `
### Exercise 5-21
```{r}
p_winsorised_mean <- function(x, p) {
  # calculate quantile cutoffs
  x_quantiles <- c(-Inf,
                   quantile(x, c(p, 1-p), names = FALSE),
                   Inf)

  # map quantiles to x indices
  x_intervals <- findInterval(x, x_quantiles)

  x_winsorised <- ifelse(x_intervals==2,
                           x,
                               ifelse(x == 1, x_quantiles[2], x_quantiles[3]))

  return(mean(x_winsorised))
}

x <- c(8, 5, 2, 9, 7, 4, 6, 1, 3)
p <- .25

p_winsorised_mean(x, p)
```
## `split()` and `unsplit()`
+ `split()` can perform grouped aggregations
+ `split()` takes a vector `x`, a factor `f`, and returns a list where each element corresponds to levels of 'f' containing evenly elements of 'x' at corresponding indices.
+ `split()` also tests our knowledge of '[' and '[['!

## `split()` and `unsplit()`

```{r}
x <- c(0, 0.2, 0.25, 0.4, 0.66, 1)

f <- findInterval(x, c(-Inf, 0.25, 0.5, 0.75, Inf))

split(x, f)
```
+ i.e. x[1:2] pairs with f[1:2], x[3:4] pairs with f[3:4]

## `split()` on data.frames
+ Use `Map()` and `split()` to on data.frames to perform grouped aggregations.

```{r}
l <- Map(
    function(x) {
        c(Min=min(x), Median=median(x), Mean=mean(x), Max=max(x))
    },
    split(iris[["Sepal.Length"]], iris[["Species"]])
)
print(l[1:3])
```

## vs. Dplyr

```{r}
iris |>
  dplyr::group_by(Species) |>
  dplyr:: summarise(Min = mean(Sepal.Length),
                    Median = median(Sepal.Length),
                    Max = max(Sepal.Length))
```
## vs. Dplyr
+ `split()` may be useful for compatibility with packages that don't have a tidyverse compatible grouping attribute
+ ???

## `order()` and `sort()`

+ `order()` returns a vector that lists *indices* corresponding to the values of 'x'
```{r}
x <- c(10, 20, 30, 50, 40)
order(x)
```

+ `sort()` returns the original vector, in ascending order
```{r}
sort(x)
```

+ `sort(x)` is equivalent to `x[order(x)]`
```{r}
all.equal(sort(x), x[order(x)])
```
## `order()` and `sort()`
+ `order()` can return rankings
```{r}
order(order(x))
# vs.
rank(x, ties = "random")
```

## `duplicated()`
+ `duplicated()` returns a logical vector indicating whether a each element in `x`  is repeated in `x`
```{r}
x <- sample(1:10, 10, replace = TRUE)
duplicated(x)
# vs.
match(x, unique(x))
```

## `tabulate()`
+ `tabulate()` returns a count of each element in vector `x`
```{r}
x <- c(2, 4, 6, 2, 2, 2, 3, 6, 6, 3)
tabulate(x)
```

## `tabulate()`
+ (Exercise 5-15)
```{r}
y <- c("a", "b", "a", "c", "a", "d", "e", "e", "g", "g", "c", "c", "g")
result <- structure(tabulate(factor(y)), # y must be a factor or numeric
                    names = letters[match(unique(y), letters)])
print(result)
```
## Subsetting and Attributes

+ Subsetting functions tend to drop all attributes except for names
+ `'[['` drops *all* attributes

## Applications
+ Replace `NA` values with values from cubic interpolation
+ (Exercise 5-23)
```{r}
euraud <- scan("https://raw.githubusercontent.com/gagolews/teaching-data/master/marek/euraud-20200101-20200630.csv", comment.char = "#")

# create index of NA values
na_index <- is.na(euraud) == TRUE

# assign interpolated values from spline()
# to the NA values using the index
euraud[na_index] <- spline(euraud, n = length(euraud)) |>
                    _[["x"]] |> # spline() returns a named list
                    _[na_index]
```

## Applications
+ Reorder a list by it's names attribute
+ (Exercise 5-14)

```{r}
x <- list(a=1, b=2)
y <- list(c=3, a=4)

union_list <- function(x,y){
# combine lists
z <- c(x,y)
# order by names (increasing) then by value (decreasing)
z <- z[order(names(z), unlist(z), 
             decreasing = c(FALSE, TRUE),
             method="radix")]
# remove duplicated names
z[!(duplicated(names(z)) == TRUE)]
}

str(union_list(x,y))

```
## Applications
+ Spearman rank coefficient
```{r}
spearman_rank <- function(x, y) {
  n <- ifelse(length(x) == length(y), length(x),
                                  warning("x and y have different lengths"))
  r_i = rank(x)
  s_i = rank(y)
  d_i = r_i - s_i

  (6 * sum(d_i^2))/
    (n*(n^2-1))
}

x <- rnorm(n = 5)
y <- rnorm(n = 5)

spearman_rank(x,y)
```
## Exercises?

+ Which exercises were the most challenging?

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
