# Matrices and other arrays

**Learning objectives: all about matrix/array**

- What is a matrix and an array (for R)

- How to create and/or convert to it

- Matrix and array manipulations: 

    - Indexing 
    
    - Common operations 
    
- Matrix algebra

## What is a matrix and an array {-}

### S3 all the way!

Give a generic vector a `dim` attributes -> matrix 

```{r}
(A <- structure(1:6, dim = c(2, 3)))
```

```{r}
class(A)
```

Ok, but an array?: 

```{r}
(B <- structure(1:6, dim=6))
class(B)
```

A matrix is a 2D array (which is a vector). The convention seems to be: give them a capital letter.

### Not just numeric

Logical:

```{r}
A >= 3
```

Character:

```{r}
matrix(strrep(LETTERS[1:6], 1:6), ncol=3)
```

Even list:

```{r}
matrix(list(1, 11:21, "A", list(1, 2, 3)), nrow=2)
```

> Certain elements are not displayed correctly, but they are still there.

Internally elements are in column-major:

```{r}
(B <- matrix(1:6, ncol=3, byrow=TRUE))
```

```{r}
as.numeric(B)
```
 cool tip: you can name dimnames!
 
```{r}
A <- structure(1:6, dim=c(2, 3), 
               dimnames=list(letters[1:2], LETTERS[1:3]))
names(dimnames(A)) <- c("ROWS", "COLUMNS")
print(A)
```
 

## How to create and/or convert to it {-}

They can be create by `array()` or by `matrix()`: 

  - `array(vector, dim = length(data), dimnames = NULL)`
  
  `dimnames` is optional but will be used later. 
  
  - `matrix(vector, nrow = 1, ncol = 1, byrow = FALSE, dimnames = NULL)`
  
  - `cbind` and `rbind` can also be used (sometimes in conjunction with `do.call()`)
  
  *Be careful* `do.call` will use the recycle rule.  
  
  - `simplify2array()`: has the name suggest: list of vector of same length  
-> `sapply()` is `lapply()` + (if `simplify = TRUE`, the default) `simplif2array`

```{r}
x <- list(a = runif(10), b = rnorm(15))
sapply(x, function(e) c(Min = min(e), Max = max(e)))
```

- `table()` creates contingency table: matrix/array 

## Matrix and array manipulations {-}

### what you know on indexing on data.frame works here

Get the book example:

```{r}
(A <- matrix(1:12, byrow=TRUE, nrow=3))
(B <- `dimnames<-`(A, list(  
    c("a", "b", "c"),      
    c("x", "y", "z", "w") 
)))
```

```{r}
A[3, ] # rows
```
```{r}
A[, 2] # column
```

```{r}
B["c", "y"]  # specify dimnames
```

**By** default `drop` is set to TRUE with `[`

```{r}
A[1, 2, drop = FALSE]
```

### But behind the scene it is just a vector:

This works! :mind_blow:

```{r}
A[6]
```

### Selecting by submatrices

It works, not going deep into it, but selecting on two columns numeric matrices is probably super useful. 

It uses an index on row/column and return individual cells matching those "coordinates".

On the same idea:

```{r}
which(A>7, arr.ind=TRUE) #
```

Return a "coordinate" matrix of cell that are matching a logical vector.

### Higher-dimensional array (mostly know as contingency table)

```{r}
str(dimnames(Titanic))
```

hence (because it is "dimnmamed") it works:

```{r}
Titanic["1st", "Male", "Adult", "No", drop = FALSE]
```

But also with empty "space":

```{r}
Titanic["1st", , "Adult", ] 
```


## "Common operations":

- You can use `t()`

- Do aggregate with `apply()` (see data.frame)

The second argument is specifying the dimensions mostly: 1 for row and 2 for columns but it can be "expanded".

```{r}
apply(Titanic, c(1, 3), mean)
```


### Binary operators

- array x array: do an element wise operation 

- array x scalar: scalar is recycled

- array x vector: the only tricky one it works by recycling column wisely

```{r}
(B <- matrix(1:6, byrow=TRUE, nrow=2))
```

```{r}
B * c(1, 100)
```


## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
